---
title: "In-Class Exercise 6"
description: |
    In this hands-on exercise, I will derive spatial weights by using functions provided by spdep package and appliy these spatial weights to compute geographically referenced attributes.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Useful Reference

[R-spatial](https://rspatial.github.io/raster/reference/index.html)

- poly2nb() is to create proximity spatial weight
- Knearneigh() to construct distance matrix and define neighbour 
  + Fixed Kernel
  + Adaptive Kernel 
- nb2listw() to convert so spatial weight

In this hands-on exercise, we will compute spatial weights using R

- import geospatial 

## Getting Started

In this hands-on exercise, the following r packages will be used :

- import geospatial data using appropriate function(s) of rgdal package,
- import csv file using appropriate function of readr package,
- perform relational join using appropriate join function of dplyr package,
- compute spatial weights using appropriate R functions, and
- calculate spatially lag attributes using appropriate functions of spdep package.

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Loading Data Into R 

### Import shapefile into r environment

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

There is a total of 29 variables 

## Performing relational join

The code chunk below will be used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

```{r}
hunan <- left_join(hunan,hunan2012)
```

Now there are 36 variables 

## Visualising Regional Development Indicator

Now, we are going to prepare a basemap and choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmap package.

```{r}
basemap <-tm_shape(hunan) + tm_polygons() + tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap,gdppc,asp=1,ncol=2)
```
