---
title: "In-Class Exercise 6"
description: |
    In this hands-on exercise, I will derive spatial weights by using functions provided by spdep package and appliy these spatial weights to compute geographically referenced attributes.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Useful Reference

[R-spatial](https://rspatial.github.io/raster/reference/index.html)

- poly2nb() is to create proximity spatial weight
- Knearneigh() to construct distance matrix and define neighbour 
  + Fixed Kernel
  + Adaptive Kernel 
- nb2listw() to convert so spatial weight

In this hands-on exercise, we will compute spatial weights using R

- import geospatial 

## Getting Started

In this hands-on exercise, the following r packages will be used :

- import geospatial data using appropriate function(s) of rgdal package,
- import csv file using appropriate function of readr package,
- perform relational join using appropriate join function of dplyr package,
- compute spatial weights using appropriate R functions, and
- calculate spatially lag attributes using appropriate functions of spdep package.

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Loading Data Into R 

### Import shapefile into r environment

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

There is a total of 29 variables 

## Performing relational join

The code chunk below will be used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

```{r}
hunan <- left_join(hunan,hunan2012)
```

Now there are 36 variables 

## Visualising Regional Development Indicator

Now, we are going to prepare a basemap and choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmap package.

```{r fig.width=26, fig.height=15}
basemap <-tm_shape(hunan) + tm_polygons() + tm_text("NAME_3", size=1)

gdppc <- qtm(hunan, "GDPPC")+
tm_layout(legend.outside = FALSE,legend.title.size = 2.10, legend.text.size = 1.90)
tmap_arrange(basemap,gdppc,asp=1,ncol=2)
```

## Computing Contiguity Spatial Weights

### Creating (ROOK) contiguity based neighbours

The code chunk below is used to compute Rook contiguity weight matrix.

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

The summary report above shows that there are 88 area units in Hunan. The most connect area unit has 10 neighbours. There are two area units with only one heighbours.



### Computing (QUEEN) contiguity based neighbours


The code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

We can retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

To reveal the count names of the five neighboring polygons, the code chunk will be used:

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

We can retrieve the GDPPC of these five counties by using the code chunk below
```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

## Plotting Hisogram to visualize distribution 

```{r}
hist(hunan$GDPPC)
```

From the output we can see that it is right skewed 


## Visualising contiguity weights

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

We do the same for latitude with one key difference. We access the second value per each centroid with [[2]].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```


Use cbind to put longitude and latitude into the same object.

```{r}
coords <- cbind(longitude, latitude)
```

Plotting Queen contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```
Plotting Rook contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")


```

## Plotting Rook contiguity based neighbours maps


```{r}


par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch=19, cex=0.6, add=TRUE, col="red",main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch=19, cex=0.6, add=TRUE, col="red",main="Rook Contiguity")

```

## Computing distance based neighbours

Note: Direct distance between centroid of one location to centroid of another
location 

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is *61.79 km*, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

## Computing fixed distance weight matrix

Now, we will compute the distance weight matrix by using dnearneigh() as shown in the code chunk below.

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```


## Computing distance based neighbours


### Determine the cut-off distance

Firstly, we need to determine the upper limit for distance band by using the steps below:

+ Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using knearneigh() of spdep.

+ Convert the knn object returned by knearneigh() into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using knn2nb().

+ Return the length of neighbour relationship edges by using nbdists() of spdep. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.

+ Remove the list structure of the returned object by using unlist().

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

## Computing fixed distance weight matrix

Now, we will compute the distance weight matrix by using dnearneigh() as shown in the code chunk below.

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

Average number of links: 3.681818” shown above means about 4 neighbours


Display the structure of the weight matrix by combining table() and card() of spdep.

```{r}
table(hunan$County, card(wm_d62))
```

## Plotting fixed distance weight matrix

Next, we will plot the distance weight matrix by using the code chunk below.

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Alternatively, we can plot both of them next to each other by using the code chunk below.

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

Note: Matrix is binary, if it is within distance it will be indicated as 1 else 0

## Computing adaptive distance weight matrix

Note: Certain location entity may be close while in other location, entity may be far apart 

Look for the 6 closest neighbour 

Control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

## Plotting distance based neighbours

We can plot the weight matrix using the code chunk below.


```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW


First, we will compute the distances between areas by using nbdists() of spdep.

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row-standardised weights matrix

Next, we need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style=“W”). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. While this is the most intuitive way to summaries the neighbors’ values it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we’ll stick with the style=“W” option for simplicity’s sake but note that other more robust options are available, notably style=“B”.

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

To see the weight of the first polygon’s four neighbors type:

```{r}
rswm_q$weights[10]
```

Each neighbor is assigned a 0.2 of the total weight. This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied.

Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.

```{r}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```
Note: Zero.policy - Exclude polygon with no neighbour e.g island 

To see the weight of the first polygon’s four neighbors type:

```{r}
rswm_q$weights[10]
```

## Application of Spatial Weight Matrix

### Spatial lag with row-standardized weights

Finally, we’ll compute the average neighbor GDPPC value for each polygon. These values are often referred to as spatially lagged values.

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

Note: Sum up and divide by 5 to know neighbour average compared to own value 


We can append the spatially lag GDPPC values onto hunan SpatialPolygonDataFrame by using the code chunk below.

```{r}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

Next, we will plot both the GDPPC and spatial lag GDPPC for comparison using the code chunk below.

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)

```
GDP per capita is much higher than neighbour in thesout eastern region of Hunan

Average of neighbour without including itself

### Spatial window sum

The spatial window sum uses and includes the diagonal element. To begin, we will assign knn6 to a new variable because we will directly alter its structure to add the diagonal elements

```{r}
knn6a <- knn6

```

To add the diagonal element to the neighbour list, we just need to use include.self() from spdep.

```{r}
include.self(knn6a)

```

Next, we will assign binary weights to the neighbour structure that includes the diagonal element.

```{r}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

Again, we use nb2listw() and glist() to explicitly assign weight values.

```{r}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
```


With our new weight structure, we can compute the lag variable with lag.listw().

```{r}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
```


Next, we will convert the lag variable listw object into a data.frame by using as.data.frame().

```{r}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
```

Note: The third command line on the code chunk above renames the field names of lag_knn6.res object into NAME_3 and lag_sum GDPPC respectively.

Next, the code chunk below will be used to append lag_sum GDPPC values onto hunan sf data.frame by using left_join() of dplyr package.

```{r}
hunan <- left_join(hunan, lag_knn6.res)
```

Lastly, qtm() of tmap package is used to plot the GDPPC and lag_sum GDPPC map next to each other for quick comparison.

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```
