---
title: "In-Class Exercise 6"
description: |
    In this hands-on exercise, I will derive spatial weights by using functions provided by spdep package and appliy these spatial weights to compute geographically referenced attributes.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-20-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Useful Reference

[R-spatial](https://rspatial.github.io/raster/reference/index.html)

- poly2nb() is to create proximity spatial weight
- Knearneigh() to construct distance matrix and define neighbour 
  + Fixed Kernel
  + Adaptive Kernel 
- nb2listw() to convert so spatial weight

In this hands-on exercise, we will compute spatial weights using R

- import geospatial 

## Getting Started

In this hands-on exercise, the following r packages will be used :

- import geospatial data using appropriate function(s) of rgdal package,
- import csv file using appropriate function of readr package,
- perform relational join using appropriate join function of dplyr package,
- compute spatial weights using appropriate R functions, and
- calculate spatially lag attributes using appropriate functions of spdep package.

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Loading Data Into R 

### Import shapefile into r environment

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### Import csv file into r environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```

There is a total of 29 variables 

## Performing relational join

The code chunk below will be used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

```{r}
hunan <- left_join(hunan,hunan2012)
```

Now there are 36 variables 

## Visualising Regional Development Indicator

Now, we are going to prepare a basemap and choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmap package.

```{r fig.width=26, fig.height=15}
basemap <-tm_shape(hunan) + tm_polygons() + tm_text("NAME_3", size=1)

gdppc <- qtm(hunan, "GDPPC")+
tm_layout(legend.outside = FALSE,legend.title.size = 2.10, legend.text.size = 1.90)
tmap_arrange(basemap,gdppc,asp=1,ncol=2)
```

## Computing Contiguity Spatial Weights

### Computing (QUEEN) contiguity based neighbours


The code chunk below is used to compute Queen contiguity weight matrix.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

Polygon 1 has 5 neighbors. The numbers represent the polygon IDs as stored in hunan SpatialPolygonsDataFrame class.

We can retrive the county name of Polygon ID=1 by using the code chunk below:

```{r}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

To reveal the count names of the five neighboring polygons, the code chunk will be used:

```{r}
hunan$NAME_3[c(2,3,4,57,85)]
```

We can retrieve the GDPPC of these five counties by using the code chunk below
```{r}
nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1
```

## Plotting Hisogram to visualize distribution 

```{r}
hist(hunan$GDPPC)
```

From the output we can see that it is right skewed 


## Visualising contiguity weights

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

We do the same for latitude with one key difference. We access the second value per each centroid with [[2]].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```


Use cbind to put longitude and latitude into the same object.

```{r}
coords <- cbind(longitude, latitude)
```

Plotting Queen contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```
Plotting Rook contiguity based neighbours maps

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

