---
title: "Hands-on Exercise 4"
description: |
  In this exercise, we will use spatstat package to perform spatial point patterns analys
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-04-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1.0 Objective
Using appropriate functions of spatstat package to perform spatial point patterns analysis

## 2.0 Getting Started
In this hands-on exercise, 5 R packages will be used:

- sf
- spatstat
- raster
- maptools
- tmap


## 2.0 Installing and Loading the R packages

This code chunk will be used to install and load these packages in RStudio.

```{r echo=TRUE, eval=TRUE}
packages = c('maptools', 'sf', 'raster','spatstat', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## 3.0 Spatial Data Wrangling 

### 3.1 Importing the spatial data

Importing the child care service geojson data

```{r echo=TRUE, eval=TRUE}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```
Importing the Costal Outline Data

```{r echo=TRUE, eval=TRUE}
sg_sf <- st_read(dsn = "data/CostalOutline.shp")

```

Importing the Subzone Data

```{r echo=TRUE, eval=TRUE}

mpsz_sf <- st_read(dsn = "data", 
                layer = "MP14_SUBZONE_WEB_PL")
```

### 3.2 Transforming & Assigning EPSG code to a simple feature data frame

```{r echo=TRUE, eval=TRUE}

st_crs(mpsz_sf) 
st_crs(sg_sf)
st_crs(childcare_sf) 
```
From the output, it can be seen that although mpsz_sf data frame is projected in svy21 but when we read until the end of the print, it indicates that the EPSG is 9001. This is a wrong EPSG code because the correct EPSG code for svy21 should be 3414.

In order to assign the correct EPSG code to mpsz_sf and sg_sf data frame, st_set_crs() of sf package is used as shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}

mpsz_sf <-st_transform(mpsz_sf, crs=3414)
sg_sf <-st_transform(sg_sf, crs=3414)

```


### 3.3 Mapping the Geospatial Data Sets

Method 1

```{r echo=TRUE, eval=TRUE}

tm_shape(sg_sf) +
  tm_polygons() +
tm_shape(mpsz_sf) +
  tm_polygons() +
tm_shape(childcare_sf)+
  tm_dots()


```

Method 2

```{r echo=TRUE, eval=TRUE}

tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()

```

Notes: Do not plot more than 10 interactive map when publishing on Netlify as
it takes up a lot of memory.

## 4.0 GeoSpatial Data Wrangling 

### 4.1 Converting sf data frames to sp’s Spatial* class

Convert the three geospatial data from simple feature data frame to SP Spatial
class.

```{r echo=TRUE, eval=TRUE}

childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
childcare
mpsz
sg
```

### 4.2 Converting the Spatial* class into generic sp format

Convert the Spatial classes into Spatial object first.

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")

childcare_sp
sg_sp
```

The foundational structure for any spatial object in sp is the Spatial class. 
It has two “slots”:

- a bounding box
- a CRS class object to define the Coordinate Reference System

Spatial objects stands for Points, Lines, or Polygons

### 4.3 Converting the generic sp format into spatstat's ppp format

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

### 4.4 Plot childcare_ppp and examine the different

```{r}
plot(childcare_ppp)
summary(childcare_ppp)
```

Notes: The statistical methodology used for spatial point patterns processes is based largely on the assumption that process are simple, that is, that the points cannot be coincident.


### 4.5 Handling duplicated points

We can check the duplication in a ppp object by using the code chunk below.
```{r}
any(duplicated(childcare_ppp))
```

To count the number of co-indicence point, we will use the multiplicity() function as shown in the code chunk below.
```{r}
multiplicity(childcare_ppp)
```
To know how many locations have more than one point event, we can use the code chunk below..
```{r}
sum(multiplicity(childcare_ppp) > 1)
```

The output shows that there are 128 duplicated point events.


To view the locations of these duplicate point events, we will plot childcare data by using the code chunk below.

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

To spot the duplicate points from the map shown: 

- Delete the duplicates but that will also mean that some useful point events will be lost.  

-  Jittering which will add a small perturbation to the duplicate points so that they do not occupy the exact same space.  

- Make each point "unique" and then attach the duplicates of the points to the patterns as **marks**, as attributes of the points. Then you would need analytical techniques that take into account these marks.

The code chunk below implements the jittering approach.

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```


```{r}
any(duplicated(childcare_ppp_jit))
```

## 5.0 Creating Owin Object

Covert SG SpatialPolygon object into owin object of spatstat.

```{r}
sg_owin <- as(sg_sp, "owin")
plot(sg_owin)
summary(sg_owin)
```


## 6.0 Combining point events object and owin object
```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
summary(childcareSG_ppp)
```

Plot the newly derived childcareSG_ppp

```{r}
plot(childcareSG_ppp)
```



## 7.0 First-order Spatial Point Patterns Analysis

### 7.1 Computing kernel density estimation using automatic bandwidth selection method

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
```

The *plot()* function of Base R is then used to display the kernel density derived.

```{r}
plot(kde_childcareSG_bw)
```

Notes: The density values of the output range from 0 to 0.000035 which is way too small to comprehend. This is because the default unit of measurement of svy21 is in meter. As a result, the density values computed is in “number of points per square meter”.


Retrieve the bandwidth used to compute the kde layer by using the code chunk below.

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

### 7.2 Rescalling KDE values

Rescale() is used to covert the unit of measurement from meter to kilometer. 

```{r}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

density() use the resale data set and plot the output kde map.

```{r}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

Output image looks identical to the earlier version, the only changes in the data values (refer to the legend).

### 7.3 Working with different automatic badwidth methods

Beside bw.diggle(), there are three other **spatstat** functions can be used to determine the bandwidth, they are:



```{r}
 bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

Compare the output of using bw.diggle and bw.ppl methods.

```{r fig.width=16, fig.height=6}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```


### 7.4 Working with different kernel methods

The code chunk below will be used to compute three more kernel density estimations by using these three kernel function.

```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

## 8.0 Fixed and Adaptive KDE

### 8.1 Computing KDE by using fixed bandwidth

```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```


### 8.2 Computing KDE by using Adaptive bandwidth

Fixed bandwidth method is very sensitive to highly skew distribution of 
spatial point patterns over geographical units for example urban versus rural. 
One way to overcome this problem is by using adaptive bandwidth instead.

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)

```

We can compare the fixed and adaptive kernel density estimation outputs by using the code chunk below. 

```{r fig.width=16, fig.height=6}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```


## 9.0 Converting KDE output into grid object.

Convert it so that it is suitable for mapping purposes

```{r}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

### 9.1  Converting gridded output into raster

Next, we will convert the gridded kernal density objects into RasterLayer object by using raster() of raster package.

```{r}

kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

Let us take a look at the properties of kde_childcareSG_bw_raster RasterLayer.  

```{r}
kde_childcareSG_bw_raster
```

Notice that the crs property is NA. 


### 9.2  Assigning Projection System

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```


### 9.3  Visualising the output

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

## 10.0 Comparing Spatial Point Patterns using KDE

### 10.1 Extract Study Area

The code chunk below will be used to extract the target planning areas.

```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```


Notes:raster values are encoded explicitly onto the raster pixel using the values in "v"" field.

Plotting target planning areas

```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

### 10.2 Converting the spatial point data frame into generic sp format

Next, we will convert these SpatialPolygonsDataFrame layers into generic spatialpolygons layers. 

```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```


### 10.3 Creating owin object

Now, we will convert these SpatialPolygons objects into owin objects that is required by spatstat.

```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

### 10.4 Combining childcare points and the study area

By using the code chunk below, we are able to extract childcare that is within the specific region to do our analysis later on.

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

Next, rescale() function is used to trasnform the unit of measurement from metre to kilometre. 

```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

The code chunk below is used to plot these four study areas and the locations of the childcare centres. 

```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```


### 10.5 Computing KDE

The code chunk below will be used to compute the KDE of these four planning area. bw.diggle method is used to derive the bandwidth of each



```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

### 10.6 Computing fixed bandwidth KDE

For comparison purposes, we will use 250m as the bandwidth.

```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```


## 11. Nearest Neighbour Analysis

Clark-Evans test of aggregation for a spatial point pattern

The test hypotheses are:

Ho = The distribution of childcare services are randomly distributed.

H1= The distribution of childcare services are not randomly distributed.

The 95% confident interval will be used.

### 11.1 Testing spatial point patterns using Clark and Evans Test

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```
Since p-vlaue is 0.01 which is less than alpha of 0.05, we will reject the null
hypothesis and conclude that the distribution of childcare services are not randomly distributed.

### 11.2 Clark and Evans Test: Choa Chu Kang planning area


Perform Clark-Evans test of aggregation for childcare centre in Choa Chu Kang planning area
```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

### 11.3 Clark and Evans Test: Tampines planning area

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```


## 12. Second-order Spatial Point Patterns Analysis

### 12. 1 Analysing Spatial Point Process Using G-Function

#### 12.1.1 Choa Chu Kang planning area

##### 12.1.1.1 Computing G-function estimation

The code chunk below is used to compute G-function using *Gest()* of **spatat** package.

```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```


##### 12.1.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang  are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with G-fucntion

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
plot(G_CK.csr)
```

#### 12.1.2 Tampines planning area


##### 12.1.2 .1  Computing G-function estimation

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

##### 12.1.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected is p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
plot(G_tm.csr)
```

## 13 Analysing Spatial Point Process Using F-Function


### 13.1 Choa Chu Kang planning area

#### 13.1.1 Computing F-function estimation

The code chunk below is used to compute F-function using *Fest()* of **spatat** package.

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

#### 13.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang  are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

Monte Carlo test with F-fucntion

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
plot(F_CK.csr)
```

### 13.2 Tampines planning area

#### 13.2.1  Computing F-function estimation

Monte Carlo test with F-fucntion

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

## 14 Analysing Spatial Point Process Using K-Function 

### 14.1 Choa Chu Kang planning area

#### 14.1.1 Computing K-fucntion estimate

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### 14.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)

```

```{r fig.width=8, fig.height=5}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### 13.3 Tampines planning area

#### 13.3.1  Computing K-fucntion estimation

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```


#### 13.3.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)

```

```{r fig.width=7,  fig.height=4.5}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## 14. Analysing Spatial Point Process Using L-Function

### 14.1 Choa Chu Kang planning area

#### 14.1.1 Computing L Fucntion estimation

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### 14.1.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Choa Chu Kang are randomly distributed.

H1= The distribution of childcare services at Choa Chu Kang are not randomly distributed.

The null hypothesis will be rejected if p-value if smaller than alpha value of 0.001.

The code chunk below is used to perform the hypothesis testing.

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)

```

```{r fig.width=8, fig.height=5}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### 14.2 Choa Chu Kang planning area

Tampines planning area

#### 14.2.1 Computing L-fucntion estimate

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```


####$ 14.2.2 Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted.  The hypothesis and test are as follows:

Ho = The distribution of childcare services at Tampines are randomly distributed.

H1= The distribution of childcare services at Tampines are not randomly distributed.

The null hypothesis will be rejected if p-value is smaller than alpha value of 0.001.

The code chunk below will be used to perform the hypothesis testing.

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)

```

Then, plot the model output by using the code chun below.

```{r fig.width=7,  fig.height=4.5}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```


