---
title: "In-class Exercise 4"
description: |
  A short description of the post.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-06-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## 1. Onemap SG API

### 1.1 Getting Started 

Installing and importing Onemap SG API.

```{r echo=TRUE, eval=TRUE}
packages <- c('onemapsgapi')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

### 1.2 Query 

```{r}
#get_theme("TOKEN", "hotels")
```

Note: Need to put double quote for token and query name 


## 2. In Class Exercise 4

### 2.1 Getting Started 

In this in-class exercise, we will be using the following library:

- SF: Import geospatial data
- Tidyverse: To read rds and to do data manipulation
- Tmap: Prepare map 
- Spatstat: Perform spatial point pattern analysis
- Maptools: Convert spatial data from Spatial Object Class to Spatstat format 
- Raster: Handle raster data


```{r echo=TRUE, eval=TRUE}
packages <- c('maptools','sf','raster','spatstat','tmap','tidyverse')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

### 2.1 Getting Started 
```{r echo=TRUE, eval=TRUE}
sg_sf <- st_read(dsn="data/shapefile", layer="CostalOutline")

mpsz_sf <- st_read(dsn="data/shapefile",layer="MP14_SUBZONE_WEB_PL")
```


### 2.2 Importing Geospatial Data
```{r echo=TRUE, eval=TRUE}
sg_sf <- st_read(dsn="data/shapefile", layer="CostalOutline")

mpsz_sf <- st_read(dsn="data/shapefile",layer="MP14_SUBZONE_WEB_PL")
```

### 2.3 Importing Aspatial Data

```{r echo=TRUE, eval=TRUE}
childcare <- read_rds("C:/Users/User/Desktop/IS415/lye-jia-wei/IS415_blog/_posts/2021-09-06-in-class-exercise-4/data/rds/childcare.rds")
CHAS <- read_rds("C:/Users/User/Desktop/IS415/lye-jia-wei/IS415_blog/_posts/2021-09-06-in-class-exercise-4/data/rds/CHAS.rds")
```

### 2.4 Converting from aspatial to geospatial

```{r echo=TRUE, eval=TRUE}
CHAS_sf <- st_as_sf(CHAS, coords = c("X_COORDINATE","Y_COORDINATE"), crs=3414)

```

```{r echo=TRUE, eval=TRUE}
childcare$Lat <- as.numeric(childcare$Lat)
childcare$Lng <- as.numeric(childcare$Lng)
childcare_sf <- st_as_sf(childcare, coords= c("Lng","Lat"),crs=4326) %>% st_transform(crs=3414)
```


## 3.0 Geospatial Data Wrangling

### 3.1 Converting from sf to Spatial classes

```{r echo=TRUE, eval=TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

### 3.2 Converting from Spatial classes to SP format

```{r echo=TRUE, eval=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
sg_sp <- as(mpsz, "SpatialPolygons")
```

