---
title: "In-class Exercise 3"
description: |
  This in-class exercise aims to identify planning subzones with extreme high or
  low numbers of aged population.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1.0 Problem Statement 
How to identify planning subzones with extreme high or low numbers of aged population?


## 2.0 Getting Started
In this in-class exercise, the key R package use is tmap package in R. Beside
tmap package, three other R packages will be used. They are:

- readr for importing delimited text file,
- tidyr for tidying data,
- sf for handling geospatial data.


This code chunk will be used to install and load these packages in RStudio.

```{r echo=TRUE, eval=TRUE}
packages <- c('sf','tidyverse','tmap')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

## 3.0 Importing Data into R 

### 3.1 Data
Two data set will be used to create percentile map and they are:

- Master Plan 2014 Subzone Boundary (Web) in ESRI shapefile
- Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of 
Dwelling, June 2011-2020 in csv format 


### 3.2 Importing Geospatial Data into R

The code chunk below uses the st_read() function of sf package to import 
MP14_SUBZONE_WEB_PL shapefile into R as a simple feature data frame called mpsz.

```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

### 3.3 Importing Attribute Data into R

The task will be performed by using read_csv() function of readr package as 
shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}
popdata <- read_csv("data/aspatial/respopagesexfa2011to2020.csv")
```

### 3.4 Data Preparation

Data table will consist of the following variables:
-YOUNG: age group 0 to 4 until age groyup 20 to 24,
-ECONOMY ACTIVE: age group 25-29 until age group 60-64,
-AGED: age group 65 and above,
-TOTAL: all age group, and
-DEPENDENCY: the ratio between young and aged against economy active group


#### 3.4.1 Data Wrangling

The following data wrangling and transformation functions will be used:

- pivot_wider() of tidyr package, and
- mutate(), filter(), group_by() and select() of dplyr package

```{r echo=TRUE, eval=TRUE}

popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

```

#### 3.4.2 Joining the Attribute Data and Geospatial Data

Convert the values in PA and SZ fields to uppercase since the values of PA and 
SZ fields are made up of upper- and lowercase. On the other, hand the SUBZONE_N 
and PLN_AREA_N are in uppercase. 

```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Join the geographical data and attribute table using planning subzone name - 
SUBZONE_N and SZ as the common identifier.

```{r echo=TRUE, eval=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

### 4.0 Analytic Mapping

#### 4.1 Boxplot

Visualisation technique to show the statistics and distribution of data values

```{r echo=TRUE, eval=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```   
From this boxplot, it can be seen that there are 6 outliers

#### 4.2 Percentile Map

#### 4.2.1  Exclude records with NA by using the code chunl below
```{r echo=TRUE, eval=TRUE}
mpsz_pop2020a <- mpsz_pop2020 %>% drop_na()
```  

#### 4.2.2 Create Quantile
```{r echo=TRUE, eval=TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- mpsz_pop2020a["DEPENDENCY"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```  
#### 4.2.3 Creating the get.var function

The code chunk is an R function to extract a variable as a vector out of an sf 
data frame.

```{r echo=TRUE, eval=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```  

#### 4.2.4 Plotting a percentile map using tmap functions

```{r echo=TRUE, eval=TRUE}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- get.var("DEPENDENCY", mpsz_pop2020a)
bperc <- quantile(var,percent)
tm_shape(mpsz_pop2020) +
  tm_polygons() +
tm_shape(mpsz_pop2020a) +
  tm_fill("DEPENDENCY",
          title="DEPENDENCY",
          breaks=bperc,
          palette="Blues",
          labels=c("< 1%", "1% - 10%",
                   "10% - 50%", 
                   "50% - 90%",
                   "90% - 99%", 
                   "> 99%"))  +
  tm_borders() +
  tm_layout(title = "Percentile Map", 
            title.position = c("right",
                               "bottom"))
```  

#### 4.2.5 Creating a percentile mapping function

The code chunk is an R function to extract a variable as a vector out of an sf 
data frame.

```{r echo=TRUE, eval=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam,df)
  bperc <- quantile(var,percent)
  tm_shape(mpsz_pop2020) +
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(title = mtitle, title.position = c("right","bottom"))
}
```  

#### 4.2.6 Testing the percentile mapping function


```{r echo=TRUE, eval=TRUE}
youngmap <- percentmap("YOUNG", mpsz_pop2020a)
agedmap <- percentmap("AGED", mpsz_pop2020a)

```  

#### 4.3 Box Map

#### 4.3.1 Creating the boxbreaks function

The following is a function that creating break points for a box map.

```{r echo=TRUE, eval=TRUE}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```  

#### 4.3.2 Creating the get.var function

The following is a function that xtract a variable as a vector out of an sf 
data frame.

```{r echo=TRUE, eval=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```  

#### 4.3.3 Testing the function

Exclude AGED = NA by using the code chunk below.

```{r echo=TRUE, eval=TRUE}
mpsz_pop2020a <- mpsz_pop2020 %>%
  filter(AGED>=0)
var <- get.var("AGED", mpsz_pop2020a)
boxbreaks(var)
```  

#### 4.3.4 Boxmap Function

```{r echo=TRUE, eval=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(mpsz_pop2020) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```  

#### 4.3.5 Using The Boxmap Function

```{r echo=TRUE, eval=TRUE}
boxmap("ECONOMY ACTIVE", mpsz_pop2020a)
```  

#### 4.4 Raw Rate Map

```{r echo=TRUE, eval=TRUE}
mpsz_pop2020a <- mpsz_pop2020a %>%
  mutate(`AGED%` = (`AGED`
/`TOTAL`)*100) %>%
  filter(`AGED%` >= 0)


var <- get.var("AGED%", mpsz_pop2020a)
boxbreaks(var)
boxmap("AGED",mpsz_pop2020a)
```  