---
title: "In-class Exercise 3"
description: |
  In this hands-on exercise, I learn how to plot functional and truthful choropleth maps by using tmap.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Getting Started

This code chunk performs three task:

- create a list of R packages needed.

```{r echo=TRUE, eval=TRUE}
packages <- c('sf','tidyverse','tmap')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

## 2. Importing Geospatial Data into R
### 2.1 Data

- Master Plan 2014 Subzone Boundary(Web)
-Singapore Redients by Planning Area/Subzone,Age Group,Sex and Type of Dwelling June 2011-2020

### 2.2 Importing Geospatial Data into R
```{r echo=TRUE, eval=TRUE}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
mpsz
```

### 2.3 Importing Attribute Data into R
### 2.2 Importing Geospatial Data into R
```{r echo=TRUE, eval=TRUE}
popdata <- read_csv("data/aspatial/respopagesexfa2011to2020.csv")
```

### 2.4 Data Preparation

Data table will consist of the following variables:
-YOUNG: age group 0 to 4 until age groyup 20 to 24,
-ECONOMY ACTIVE: age group 25-29 until age group 60-64,
-AGED: age group 65 and above,
-TOTAL: all age group, and
-DEPENDENCY: the ratio between young and aged against economy active group


#### 2.4.1 Data Wrangling
```{r echo=TRUE, eval=TRUE}

popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

```

#### 2.4.2 Joining the Attribute Data and Geospatial Data

```{r echo=TRUE, eval=TRUE}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = funs(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```
Convert the values in PA and SZ fields to uppercase since the values of PA and SZ fields are made up of upper- and lowercase. 

```{r echo=TRUE, eval=TRUE}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```
Join the geographical data and attribute table using planning subzone name - SUBZONE_N and SZ as the common identifier.

## 3. Chloropleth Mapping Geospatial Data

### 3.1 Plotting choropleth map using qtm()
```{r echo=TRUE, eval=TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

### 3.2 Creating a choropleth map by using tmap’s elements()
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

#### 3.2.1 Drawing Base Map
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 3.2.2 Drawing Choropleth Map using TM_Polygons()
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

#### 3.2.3 Drawing Choropleth Map using TM_Fill() and TM_Border()
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Adding border
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY") +
  tm_borders(lwd = 0.1,  alpha = 1)
```

### 3.3 Data classification methods of tmap
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "jenks") +
  tm_borders(alpha = 0.5)
```
Quantile data classification that used 5 classes.

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "equal") +
  tm_borders(alpha = 0.5)
```
Equal data classification method

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "kmeans")
```
Kmeans data classification method

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "fisher")
```
fisher classification method

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 5,
          style = "hclust")
```
hclust classification method

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 20,
          style = "hclust")
```
hclust classification method with 20 classes 


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 2,
          style = "hclust")
```
hclust classification method with 2 classes 

### 3.4 Color Scheme 
#### 3.4.1 Using Colorbrewer palette

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```
Reverse shading

### 3.5 Map Layouts
#### 3.5.1 Map Legend


```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

#### 3.5.2 Map Style
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### 3.5.3 Catographic Furniture 
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))

```

### 3.6.3 Drawing Small Multiple Choropleth Maps 

####3.6.3.1 Assigning Multiple Values To At Least One Aesthetic Arguments
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")

```

```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))

```
Assigning multiple values to at least one of the aesthetic arguments

####3.6.3.2 Defining a Group-By Variable 
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

####3.6.3.2 Creating Multiple Stand-alone maps with Tmap_Arrange()
```{r echo=TRUE, eval=TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```


###3.7 Mappping Spatial Object Meeting a Selection Criterion
```{r echo=TRUE, eval=TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```