---
title: "Hands-on Exercise 5"
description: |
  Network Constrained Spatial Point Patterns Analysis (NetSPPA)
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-11-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Objective 

- To derive network constrained kernel density estimation (NetKDE)
- To perform network G-function and k-function analysis

## 2. Data 

For the purpose of this study, two geospatial data sets will be used. They are:

- Punggol_St, a line features geospatial data which store the road network within Punggol Planning Area.
- Punggol_CC, a point feature geospatial data which store the location of childcare centres within Punggol Planning Area.

Both data sets are in ESRI shapefile format.

## 3. Getting Started

In this hands-on exercise, four R packages will be used:

- spNetwork to perform Spatial Point Patterns Analysis such as kernel density estimation (KDE) and K-function on network

- rgdal to import geospatial data in R and store as sp objects.

- sp which provides classes and methods for dealing with spatial data in R.

- tmap which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using leaflet API.

Use the code chunk below to install and launch the four R packages.

```{r echo=TRUE, eval=TRUE}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## 3. Data Import & Preparation

The code chunk below uses readOGD() of rgdal package to important 
Punggol_St and Punggol_CC geospatial data sets into RStudio as 
SpatialLinesDataFrame and SpatialPointsDataFrame respectively.

```{r echo=TRUE, eval=TRUE}
network <- readOGR(dsn="data/geospatial", 
                   layer="Punggol_St",
                   verbose = FALSE)
childcare <- readOGR(dsn="data/geospatial",
                     layer="Punggol_CC",
                     verbose = FALSE)
```


Next, print the content of network SpatialLineDataFrame and childcare 
SpatialPointsDataFrame by using the code chunk below.

```{r echo=TRUE, eval=TRUE}
str(network)
str(childcare)
```

Note:  spNetwork is expecting the geospatial data contains complete CRS information.

In the code chunk below, spTransform() of sp package is used to assign EPSG code to the SpatialDataFrames. The epsg:3414 is the code for svy21.

```{r echo=TRUE, eval=TRUE}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

## 4. Visualizing the Geospatial Data

Use plot() of Base R

```{r echo=TRUE, eval=TRUE}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```


To visualise the geospatial data with high cartographic quality and interactive manner, 
the mapping function of tmap package can be used as shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()
```


## 5. Network Constrained KDE (NetKDE) Analysis


### 5.1 Preparing the lixels objects

Before computing NetKDE, the SpatialLines object need to be cut into lixels with a specified minimal distance. This task can be performed by using with lixelize_lines() of spNetwork as shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}
lixels <- lixelize_lines(network,700,mindist = 350)
```

- The length of a lixel, lx_length is set to 700m, and
- The minimum length of a lixel, mindist is set to 350m.

### 5.2 Generating line centre points

Next, lines_center() of spNetwork will be used to generate a SpatialPointsDataFrame (i.e. samples) with line centre points as shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}
samples <- lines_center(lixels)
```



### 5.3 Performing NetKDE

We are ready to computer the NetKDE by using the code chunk below.

```{r echo=TRUE, eval=TRUE}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

Note: Possible kernel methods supported by spNetwork are: triangle, gaussian, scaled gaussian, tricube, cosine ,triweight, epanechnikov or uniform.
 
Currently, spNetwork support three popular methods - simple, discontinuous, continuous 

### 5.4 Visualising NetKDE

The code chunk below will be used to insert the computed density values (i.e. densities) into samples and lixels objects as density field.

```{r echo=TRUE, eval=TRUE}
samples$density <- densities
lixels$density <- densities
```

The code chunk below is used to resale the density values from number of events per meter to number of events per kilometer.

```{r echo=TRUE, eval=TRUE}
# rescaling to help the mapping
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```

The code below uses appropriate functions of tmap package to prepare interactive and high cartographic quality map visualisation.

```{r echo=TRUE, eval=TRUE}

tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
```

The interactive map above effectively reveals road segments (darker color) with relatively higher density of childcare centres than road segments with relatively lower density of childcare centres (lighter color)

## 6. Network Constrained G- and K-Function Analysis

Perform complete spatial randomness (CSR) test by using kfunctions() of spNetwork package. The null hypothesis is defined as:

Ho: The observed spatial point events (i.e distribution of childcare centres) are uniformly distributed over a street network in Punggol Planning Area.

The CSR test is based on the assumption of the binomial point process which implies the hypothesis that the childcare centres are randomly and independently distributed over the street network.

If this hypothesis is rejected, we may infer that the distribution of childcare centres are spatially interacting and dependent on each other; as a result, they may form nonrandom patterns.

```{r echo=TRUE, eval=TRUE}
kfun_childcare <- kfunctions(network, 
                             childcare,
                             start = 0, 
                             end = 1000, 
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)
```


Visualise the ggplot2 object of k-function by using the code chunk below.

```{r echo=TRUE, eval=TRUE}
kfun_childcare$plotk
```

The blue line is the empirical network K-function of the childcare centres in Punggol planning area. The gray envelop represents the results of the 50 simulations in the interval 2.5% - 97.5%. Because the blue line between the distance of 250m-400m are below the gray area, we can infer that the childcare centres in Punggol planning area resemble regular pattern at the distance of 250m-400m.