---
title: "Hands-on Exercise 3"
description: |
  This hands-on exercise covers how to plot functional and truthful choropleth maps using tmap package.
author:
  - name: Lye Jia Wei
    url: {}
date: 08-29-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading Libraries

- sf: handle geospatial data
- tmap: plot choropleth maps
- tidyverse: perform data science tasks (i.e. importing, wrangling and visualising data)

```{r echo=TRUE}
packages = c('tmap', 'sf', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Importing Data into R

### Import Geospatial Data

*st_read()* of sf package is used to import the geospatial data

```{r echo=TRUE}
mpsz = st_read(dsn="data/geospatial",
                layer="MP14_SUBZONE_WEB_PL")

mpsz
```

### Import Aspatial Data

*read_csv()* of readr package is used to import the CSV file

```{r echo=TRUE}
popagsex <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### Data Preparation

Before a thematic map can be prepared, it is required to prepare a data table with year 2020 values. The data table should include the variables `PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, `TOTAL`, `DEPENDENCY`.

- YOUNG: age group 0-4 until age group 20-24,
- ECONOMY ACTIVE: age group 25-29 until age group 60-64,
- AGED: age group 65 and above,
- TOTAL: all age group, and
- DEPENDENCY: the ratio between young and aged against economy active group

#### Data Wrangling

- *pivot_wider()* of tidyr package: "widens" data, increasing number of columns and decreasing the number of rows
- *mutate()* of dplyr package: adds new variables and preserves existing ones
- *filter()* of dplyr package: subsets a data frame, retaining all rows that satisfy your conditions
- *group_by()* of dplyr package: groups variables together 
- *select()* of dplyr package: selects (and optionally rename) variables in a dataframe

```{r echo=TRUE}
popagsex_2020 <- popagsex %>% filter(Time == 2020) %>%
                group_by(PA, SZ, AG) %>%
                summarise(`POP` = sum(`Pop`)) %>%
                ungroup()%>%
                pivot_wider(names_from=AG, values_from=POP) %>%
                mutate(`YOUNG`=rowSums(.[3:6])+rowSums(.[12])) %>%
                mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])
                + rowSums(.[13:15]))%>%
                mutate(`AGED`=rowSums(.[16:21])) %>%
                mutate(`TOTAL`=rowSums(.[3:21])) %>%  
                mutate(`DEPENDENCY`=(`YOUNG`+`AGED`)/`ECONOMY ACTIVE`) %>%
                select(`PA`, `SZ`, `YOUNG`, `ECONOMY ACTIVE`, `AGED`, 
                `TOTAL`, `DEPENDENCY`)
popagsex_2020
```

#### Joining Attribute Data and Geospatial Data

Before performing georelational join, it is required to convert values in PA and SZ fields to uppercase

```{r echo=TRUE}
popagsex_2020 <- popagsex_2020 %>% 
            mutate_at(.vars = vars(PA, SZ),.funs = funs(toupper)) %>%
            filter(`ECONOMY ACTIVE` > 0)
```

*left_join()* is used to join geographical data and attribute table using planning subzone name as the common identifier

```{r echo=TRUE}
mpsz_pop2020 <- left_join(mpsz, popagsex_2020,
                          by = c("SUBZONE_N" = "SZ"))
```

Note: *left_join()* is used with **mpsz** sf datframe as the left data table to ensure that the output will be a sf datframe

#### Saving Data as RDS File 

```{r echo=TRUE}
mpszpop2020 <- write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## Choropleth Mapping Geospatial Data

### Plotting a choropleth map quickly using qtm()

- Easiest and quickest way
- Conciese and provides a good default visualisation in many cases

```{r echo=TRUE}
tmap_mode("plot")
qtm(mpsz_pop2020, fill = "DEPENDENCY")
```

- *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used
- Fill argument is used to map the attribute (i.e. DEPENDENCY)

### Creating a choropleth map by using tmap's elements

Disadvantages of *qtm()*: aesthetics of individual layers are hard to control. **tmap**'s drawing elements should be used to draw a high quality cartographic choropleth map

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title="Distribution of Dependency Ratio by planning subzone",
          main.title.position = "center",
          main.title.size = 1.2,
          legend.height = 0.45, 
          legend.width = 0.35,
          frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS",
          position = c("left", "bottom"))
```

#### Drawing a base map

- Basic building block of tmap is *tm_shape()* followed by one or more layer elements such as *tm_fill()* and *tm_polygons()*
- *tm_shape()* defines the **input data** and *tm_polygons()* is used to draw the **planning subzone polygons**

```{r echo=TRUE}
tm_shape(mpsz_pop2020) + tm_polygons()
```

#### Drawing a choropleth map using tm_polygons()

To draw a choropleth showing the geographical distribution of a selected variable by planning subzone, just need to assign the target variable to *tm_polygons()*

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_polygons("DEPENDENCY")
```


- default interval binning: "pretty"
- default colour scheme: YlOrRd of ColorBrewer
- by default, missing values will be shaded in grey

#### Drawing a choropleth map using tm_fill() and tm_border()

*tm_polygons()* is a wrapper of *tm_fill()* and *tm_border()*: *tm_fill()* shades the polygons while *tm_borders()* adds borders of the shapefile onto the choropleth map

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY")
```

To add the boundary of the planning subzones, *tm_borders()* will be used:

```{r echo=TRUE}
tm_shape(mpsz_pop2020) + 
  tm_fill("DEPENDENCY") + 
  tm_borders(lwd = 0.1,  alpha = 1)
```

Arguments of *tm_borders()*:

- alpha: define transparency number between 0 (totally transparent) and 1 (not transparent, by default)
- col: border colour
- lwd: border line width, default is 1
- lty: border line type, default is “solid”


### Data classification methods of tmap

tmap provides 10 data classification methods: fixed, sd, equal, pretty (default), quantile, kmeans, hclust, bclust, fisher, and jenks

To define a data classification method, the style argument of *tm_fill()* or *tm_polygons()* will be used


#### Plotting choropleth maps with built-in classification methods

Jenks data classification that uses 5 classes

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 5, style = "jenks") +
        tm_borders(alpha = 0.5)
```

Quantile data classification 

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 5, style = "quantile") +
        tm_borders(alpha = 0.5)
```

Equal data classification with 5 classes

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 5, style = "equal") +
        tm_borders(alpha = 0.5)
```

Equal data classification with 10 classes

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 10, style = "equal") +
        tm_borders(alpha = 0.5)
```

Equal data classification with 2 classes

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 2, style = "equal") +
        tm_borders(alpha = 0.5)
```

Equal data classification with 20 classes

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ tm_fill("DEPENDENCY", n = 20, style = "equal") +
        tm_borders(alpha = 0.5)
```

#### Plotting choropleth map with custom break

For all built-in styles, category breaks are computed internally. In order to override defaults, the breakpoints can be set explicitly using the **breaks argument** of *tm_fill()*

Note: the breaks include a minimum and maximum. In order to end up with n categories, n+1 elements must be specified in the breaks option (the values **must be in increasing order**)

Good practice to get descriptive statistics on the variable before setting the break points

```{r echo=TRUE}
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, we set break points at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which is set at 0 and 100. The breaks vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00)

```{r echo=TRUE}
tm_shape(mpsz_pop2020) + tm_fill("DEPENDENCY",
                         breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
                         tm_borders(alpha = 0.5)
```


### Colour Scheme

tmap supports colour ramps either defined by the user or a set of predefined colour ramps from the RColorBrewer package

#### Using Colourbrewer Palette

To change the colour, the preferred colour is assigned to **palette argument** of *tm_fill()* as shown in the code chunk below

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          n = 6,
          style = "quantile",
          palette = "Blues") +
  tm_borders(alpha = 0.5)
```

To reverse the colour shading, add a “-” prefix

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```

### Map Layouts

- Combination of all map elements into a cohensive map
- Map elements can include: title, scale bar, compass, margins and aspect ratios


#### Map Legend

In tmap, several legend options are provided to change the placement, format and appearance of the legend

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```


#### Map Style

tmap allows a wide variety of layout settings to be changed with the help of *tmap_style()*

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### Cartographic Furniture

Beside map style, tmap also also provides arguments to draw other map furniture such as compass, scale bar and grid lines

*tm_compass()*, *tm_scale_bar()* and *tm_grid()* are used to add compass, scale bar and grid lines onto the choropleth map

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues",
          title = "No. of persons") +
  tm_layout(main.title = "Distribution of Dependency Ratio \nby planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

To reset the default style:

```{r echo=TRUE}
tmap_style("white")
```


### Drawing Small Multiple Choropleth Maps

Small multiple maps (i.e. facet maps) enable the visualisation of how spatial relationships change with respect to another variable, such as time

In tmap, small multiple maps can be plotted in 3 ways:

- by assigning multiple values to at least one of the aesthetic arguments
- by defining a group-by variable in *tm_facets()*
- by creating multiple stand-alone maps with *tmap_arrange()*

#### By assigning multiple values to at least 1 asthetic arguments

Small multiple choropleth maps are created by defining **ncols** in *tm_fill()*

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

Small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments

```{r echo=TRUE}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### By defining a group-by variable in tm_facets() 

Multiple small choropleth maps are created by using **tm_facets()**

```{r echo=TRUE}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE, 
            drop.shapes=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### By creating multiple stand-alone maps with tmap_arrange()

Multiple small choropleth maps are created by creating multiple stand-alone maps with *tmap_arrange()*

```{r echo=TRUE}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### Mapping Spatial Object Meeting a Selection Criterion

Instead of creating small multiple choropleth map, selection function can be used to map spatial objects meeting the selection criterion

```{r echo=TRUE}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```