---
title: "In-class Exercise 2"
description: |
  In this hands-on exercise, I learn how to handle geospatial data in R by using sf package.
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 08-23-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1. Getting Started

This code chunk performs three task:

- create a list of R packages needed.

```{r echo=TRUE, eval=TRUE}
packages <- c('sf','tidyverse')
for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}

```

## 2. Importing Geospatial Data

### 2.1 Importing polygon feature data in shapefile format
```{r echo=TRUE, eval=TRUE}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```
Geospatial objects are multipolygon features, with a total of 323 multipolygon features and 15 fields in mpsz simple feature data frame with svy21 projected coordinates system.

### 2.2 Importing polyline feature data in shapefile for
```{r echo=TRUE, eval=TRUE}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPath")
```
There are a total of 1625 features with svy21 projected coordinates system.

### 2.3 Importing GIS data in kml format
```{r echo=TRUE, eval=TRUE}
preschool = st_read("data/geospatial/pre-schools-location-kml.kml")
```
There a total of 1359 point features and 2 fields in wgs84 coordinates system. 

## 3. Checking the Content of Simple Feature Data Frame
```{r echo=TRUE, eval=TRUE}
st_geometry(mpsz)
glimpse(mpsz)
```
FMEL-UPD_D field is in date data type and X_ADDR, Y_ADDR, SHAPE_L and SHAPE_AREA fields are all in double-precision values.

```{r echo=TRUE, eval=TRUE}
head(mpsz, n=5)  
```

## 4. Plotting the Geospatial Data

```{r echo=TRUE, eval=TRUE}
plot(mpsz)
plot(st_geometry(mpsz))
plot(mpsz["PLN_AREA_N"])
```
## 5. Working with Projection
### 5.1 Assigning EPSG code to a simple feature data frame

```{r echo=TRUE, eval=TRUE}
st_crs(mpsz)
```
Result shows that EPSG is 9001 which is a wrong EPSG code because the correct EPSG code for svy21 should be 3414.

```{r echo=TRUE, eval=TRUE}
mpsz3414 <- st_set_crs(mpsz, 3414)
```
This is to assign the correct EPSG code to mpsz data frame

```{r echo=TRUE, eval=TRUE}
st_crs(mpsz3414)
```
Check the CSR

### 5.2 Transforming the projection of preschool from wgs84 to svy21
```{r echo=TRUE, eval=TRUE}
preschool3414 <- st_transform(preschool, 
                              crs = 3414)
```

## 6. Importing and Converting An Aspatial Data

### 6.1 Importing the aspatial data
```{r echo=TRUE, eval=TRUE}
listings <- read_csv("data/aspatial/listings.csv")
glimpse(listings) 
```

### 6.2 Creating a simple feature data frame from an aspatial data frame

```{r echo=TRUE, eval=TRUE}
listings_sf <- st_as_sf(listings, 
                       coords = c("longitude", "latitude"),
                       crs=4326) %>%
  st_transform(crs = 3414) 
```
Converts listing data frame into a simple feature data frame 

```{r echo=TRUE, eval=TRUE}
glimpse(listings_sf)
```
Examine the content of this newly created simple feature data frame.

## 7. Geoprocessing with sf package

### 7.1 Buffering

```{r echo=TRUE, eval=TRUE}
buffer_cycling <- st_buffer(cyclingpath, 
                               dist=5, nQuadSegs = 30)
```
Compute the 5-meter buffers around cycling paths

```{r echo=TRUE, eval=TRUE}
buffer_cycling$AREA <- st_area(buffer_cycling)
```
Calculating the area of the buffers

```{r echo=TRUE, eval=TRUE}
sum(buffer_cycling$AREA)
```
Find out the total land involved

### 7.2 Point-in-polygon count
```{r echo=TRUE, eval=TRUE}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```
Identify pre-schools located inside each Planning Subzone and calculate numbers of pre-schools that fall inside each planning subzone

```{r echo=TRUE, eval=TRUE}
summary(mpsz3414$`PreSch Count`)
```
Check the summary statistics of the newly derived PreSch Count field

```{r echo=TRUE, eval=TRUE}
top_n(mpsz3414, 1, `PreSch Count`)
```
List the planning subzone with the most number of pre-school

```{r echo=TRUE, eval=TRUE}
mpsz3414$Area <- mpsz3414 %>%
  st_area()

mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/ Area * 1000000)
```
Derive the area of each planning subzone and compute the density

## 8. Explorotary Data Analysis (EDA)

### 8.1 Histogram
```{r echo=TRUE, eval=TRUE}
hist(mpsz3414$`PreSch Density`)

ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
      x = "Pre-school density (per km sq)",
      y = "Frequency")
```
Histogram to reveal the distribution of PreSch Density

### 8.2 Scatterplot


```{r echo=TRUE, eval=TRUE}
ggplot(data=mpsz3414, aes(x= as.numeric(`PreSch Density`), y=`PreSch Count`)) +
  geom_point()+
labs(title = "Relationship between Pre-school Density and Pre-school Count.",
       subtitle= "change later",
      x = "Pre-school density (per km sq)",
      y = "Pre-school Count")

```