---
title: "In-class Exercise 5"
description: |
  In this in-class exercise, we will explore L-function
author:
  - name: Lye Jia Wei
    url: https://lye-jia-wei.github.io/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, eval=TRUE,echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = TRUE)
```

Important to set fig.retina to be 3 in order for subsequent visualization to be
sharp.


## Installing and Loading the R package

- maptools is for creating spatial objectes 
- sf is for importing geospatialdata 
- raster is for creating raster object
- spatstat is for performing Spatial Point Pattern Analysis (SPPA)
- tmap is for plotting thematic maps
- tidyverse is for data handling

```{r}

packages <- c('maptools','sf','raster','spatstat','tmap','tidyverse', 'plotly','ggthemes')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}
library(p, character.only=T)

```

## Import the Geospatial Data

The code chunk import shapefile using *st_read()* of **sf** package. The output object
is in tibble sf object class. 

Pgl_sf and mpsz_sf geospatial data sets into RStudio. 

Note: 
- Do not paste directory in dsn

- No extension require as this is a SHP file 

```{r}
Pgl_sf <- st_read(dsn="data/shapefile", layer="Punggol_St")
                                                                                                                                           
mpsz_sf <- st_read(dsn="data/shapefile",layer="MP14_SUBZONE_WEB_PL")
```

Projection is SVY21 

## Import aspatial data from rds folder

*read_rds()* of **readr** package is used instead of *readRDS()* of base R is used. This
is because output of *read_rds()* is in tibble object. 

Note:

- No need to set path

- Note that there are some data issue in childcare data frame with the `Lat` and `Lng`
column, the columns should be in numeric data type. The coordinate fields are
in decimal degrees. Hence, wgs referencing system is assumed. 

```{r}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

## Converting the aspatial data frame into sf objects 

```{r}
CHAS_sf <- st_as_sf(CHAS, coords = c("X_COORDINATE","Y_COORDINATE"),crs=3414)


childcare_sf <- st_as_sf(childcare, coords= c("Lng","Lat"),crs=4326) %>% st_transform(crs=3414)
```

Note:

- *st_as_sf()* accept coordinates in character data type. 

- No significant difference whether coordinate data type is change 

## Plotting for Reviewing 

Alpha is to set transparency

```{r}
tmap_mode("view")
tm_shape(childcare_sf)+tm_dots(alpha= 0.4, col="blue", size=0.05)+ tm_shape(CHAS_sf)+tm_dots(alpha= 0.4, col="red", size=0.05)
```
Note:

- childcare is aspatial, need to use childcare_sf 

## Geospatial Data Wrangling 

### Converting from sf to Spatial* data frame

as_Spatial() of sf package. 

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Converting Spatial* data frame into Spatial* Objects

as.SpatialPoint() of as.SpatialPolygon() of maptools package 

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
MPSZ_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* objects into ppp objects

This is to drop the geographical projection information using as.ppp() of
maptools package. 

```{r}

childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")

```

### Removing duplicate points using jitter 

```{r}

childcare_ppp_jit <- rjitter(childcare_ppp, retry = TRUE, nsim=1, drop = TRUE)

any(duplicated(childcare_ppp_jit))  

```

Output return a 'False' which suggest that there are no more duplicate.

Note:

- It will take some time 


```{r}

CHAS_ppp_jit <- rjitter(CHAS_ppp, retry = TRUE, nsim=1, drop = TRUE)

any(duplicated(CHAS_ppp_jit))  

```

Output return a 'False' which suggest that there are no more duplicate.

Note:

- tmap do not understand ppp for view mode


### Extracting Punggol Planning Area

Extracting planning sub zone `PUNGGOL` from data and then planning area. 

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

Note:

- Need to put a comma at the back
- Only one single object class and it belong to `PUNGGOL` planning area


### Converting Spatial Polygon Data frame into Spatial Polygon object

```{r}

pg_sp <- as(pg, "SpatialPolygons")
  
```

### Converting Spatial Polygons into owin object 

Convert into owin object which is a list of length 5. 

```{r}

pg_owin <- as(pg_sp, "owin")
  
```


### Extracting spatial points window owin

```{r}

childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]

```


```{r}
plot(childcare_pg)
```


## L-function


```{r}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim = 99,
                        rank = 1,
                        gloval = TRUE)

```


```{r}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```